Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-10-31 doc/prelim/presentation/prelim_presentation.tex
adding solution evolution pngs

\begin{algorithm}[H]
  \begin{algorithmic}[1]
    \STATE $\ve{r}_0 := \ve{b}-\ve{A}\ve{x}_0$
    \STATE $\beta := ||\ve{r}_0||_2$
    \STATE $\ve{v}_1 := \ve{r}_0 / \beta$
    \COMMENT{Create the orthonormal basis for the Krylov subspace}
    \FOR{$j = 1, 2, \cdots, m$}
    \STATE $h_{ij} \leftarrow \langle w_j,v_j \rangle$
    \STATE $w_j \leftarrow w_j - h_{ij}v_i$
    \ENDFOR
    \STATE $h_{j+1,j} \leftarrow ||w_j||_2$
    \STATE $v_{j+1} \leftarrow w_j / h_{j+1,j}$
    \COMMENT{Apply the orthogonality constraints}
    \STATE $\ve{y}_m \leftarrow argmin_y ||\beta \ve{e}_1 - \ve{H}_m\
    \ve{y}||_2 $
    \STATE $\ve{x}_m \leftarrow \ve{x}_0 + \ve{V_m} \ve{y}_m$
  \end{algorithmic}
  \caption{GMRES Iteration}
\end{algorithm}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Solution Methods for Discrete Linear Systems}

  \begin{itemize}
    \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
    \item Earliest published reference in 1950
    \item General lack of published work
    \item Modern work by Evans and others has yielded new applications
    \item No indication of parallel methods in the literature
  \end{itemize}
\end{frame}

@@

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallelization of Monte Methods}

>>>>

\begin{algorithm}[H]
  \begin{algorithmic}[1]
    \STATE $\ve{r}_0 := \ve{b}-\ve{A}\ve{x}_0$
    \STATE $\beta := ||\ve{r}_0||_2$
    \STATE $\ve{v}_1 := \ve{r}_0 / \beta$
    \COMMENT{Create the orthonormal basis for the Krylov subspace}
    \FOR{$j = 1, 2, \cdots, m$}
    \STATE $\ve{w}_j := \ve{A}\ve{v}_j$
    \FOR{$i = 1, 2, \cdots, j$}
    \STATE $h_{ij} \leftarrow \langle \ve{w}_j,\ve{v}_j \rangle$
    \STATE $\ve{w}_j \leftarrow \ve{w}_j - h_{ij}\ve{v}_i$
    \ENDFOR
    \STATE $h_{j+1,j} \leftarrow ||\ve{w}_j||_2$
    \STATE $\ve{v}_{j+1} \leftarrow \ve{w}_j / h_{j+1,j}$
    \ENDFOR
    \COMMENT{Apply the orthogonality constraints}
    \STATE $\ve{y}_m \leftarrow argmin_y ||\beta \ve{e}_1 - \ve{H}_m\
    \ve{y}||_2 $
    \STATE $\ve{x}_m \leftarrow \ve{x}_0 + \ve{V_m} \ve{y}_m$
  \end{algorithmic}
  \caption{GMRES Iteration}
\end{algorithm}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Methods for Discrete Linear Systems}

  \begin{itemize}
    \item First proposed by J. Von Neumann and S.M. Ulam in the 1940's
    \item Earliest published reference in 1950
    \item General lack of published work
    \item Modern work by Evans and others has yielded new applications
  \end{itemize}
\end{frame}

@@

  \begin{figure}[h!]
    \begin{center}
      \includegraphics[width=4in]{direct_1.png}
    \end{center}
    \caption{\textbf{Direct solution to Poisson Equation.} \textit{1
        history per state, 2500 total histories. 0.785 seconds CPU
        time} }
  \end{figure}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Direct vs. Adjoint Analysis}

  \begin{itemize}
  \item Analysis needed to select Monte Carlo method
  \item Time-dependent 2-dimensional Poisson equation
  \item Spectral radius fixed
  \item Sparsity varied with 2 Laplacian stencils
  \end{itemize}

  \[
  \nabla^2_5 = \frac{1}{\Delta^2}[u_{i-1,j} + u_{i+1,j} + u_{i,j-1} +
    u_{i,j+1} - 4 u_{i,j}]
  \]
  \[
  \begin{split}
    \nabla^2_9 = \frac{1}{6\Delta^2}[4 u_{i-1,j} + 4 u_{i+1,j} + 4
      u_{i,j-1} + 4 u_{i,j+1} + u_{i-1,j-1}\\ + u_{i-1,j+1} +
      u_{i+1,j-1} + u_{i+1,j+1} - 20 u_{i,j}]
  \end{split}
  \]

  \begin{itemize}
  \item Implicit Euler time differencing
  \end{itemize}

  \[
  \ve{A} \ve{u}^{n+1} = \ve{u}^n
  \]

%%---------------------------------------------------------------------------%%
\begin{frame}{Direct vs. Adjoint Analysis}

    \begin{column}{0.5\textwidth}
      \begin{figure}[h!]
        \centering
        \includegraphics[width=2.5in,clip]{Adjoint_Direct_CPU_Time.pdf}
        \caption{\textbf{CPU Time (s) to converge vs. Problem Size ($N$ for
            an $N \times N$ square mesh).} }
      \end{figure}
    \end{column}

    \begin{column}{0.5\textwidth}
      \begin{figure}[h!]
        \centering
        \includegraphics[width=2.5in,clip]{Adjoint_Direct_Iterations.pdf}
        \caption{\textbf{Iterations to converge vs. Problem Size ($N$ for an
            $N \times N$ square mesh).} }
      \end{figure}
    \end{column}

%%---------------------------------------------------------------------------%%
\begin{frame}{Direct vs. Adjoint Analysis}

  \begin{figure}[h!]
    \centering
    \includegraphics[width=2.5in,clip]{Adjoint_Direct_Convergence.pdf}
    \caption{\textbf{Infinity norm of the solution residual
        vs. iteration number for a problem of fixed size.} }
  \end{figure}

  \begin{itemize}
  \item CPU time dominating factor in method selection
  \item Significant speedup with adjoint method
  \item Does not affect convergence behavior
  \item Use adjoint with MCSA and Sequential Monte Carlo
  \end{itemize}

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallelization of Monte Carlo Methods}
